---
title: 'Co-epidemic of Tuberculosis Infection and Chronic Diseases'
author: ""
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2:
    toc: false
bibliography: []
csl: american-journal-of-epidemiology.csl
---

```{r library, include=FALSE, echo=FALSE, cache.lazy=FALSE}
library(table1)
library(knitr)
library(tidyverse)
library(finalfit)
library(readr) # read RDA file
library(here)
```



```{r setup, include=FALSE, echo=FALSE, cache.lazy=FALSE}

knitr::opts_chunk$set(warning=FALSE)

# Rounding rules
options(digits = 2)

# Loading results of table 1 and table 2
table1.2 <- readRDS(here("results" ,"table1.2.Rda"))
table1.1 <- readRDS(here("results","table1.1.Rda"))
table1.3 <- readRDS(here("results","table1.3.Rda"))
table2 <- readRDS(here("results", "table2.Rda"))

explanatory <- c("age","sex1","BMI","race","US_born", "marital",
 "education","income","HH_member","hypertension","diabetes","HD",     "H_TCHOL","Low_HDL","H_LDL", "H_TG","CKD", "HAV", "hepatitis")

dependent <- "TB_Infection"
```
# Authors

* Trang Quach^1^, Department of Epidemiology and Biostatistics, The University of Georgia, Athens, GA, USA

^1^ Corresponding Author: Trang Quach

Address: 101 Buck Rd, Miller Hall, Athens, Georgia 30606   

Email: thq66243@uga.edu

# Abstract


*Keywords*: Tuberculosis infection; Chronic diseases

# QUESTION FOR DR. HANDEL
In table 3 (multivariate and univariate analysis), I use the finalfit function (from finalfit package) to produce this table.

Let look at variables sex, the odds ratio is 0.75 (Male vs. Female). But when looking at the column 3 and 4 (about percentage). I realize that the OR should be >1. I double check the data. The numbers in column 3 and 4 are correct.How can I fix the OR?

Another question is about table 1. I tried tableone and finalfit package 
to produce the table 1 which is about the characteristics of participants. Both packages do not give the format tables that I want (output tables are hard to read).

I also tried another package called table1. This package provide the output table that I want. Howerver, the table is only html output and I cannot have word document. Is there any way to have this table 1 in word? In product, I include 'ask_question_for table 1' html file which has table 1 in format that I want.

Thank you.
 

# INTRODUCTION

# METHODS

## Data Collection


# RESULTS

```{r , echo=FALSE}
knitr::kable(table1.2,caption = "**Characteristics of participants**")

```







```{r , echo=FALSE}
knitr::kable(table1.3, row.names = FALSE, caption = "**Characteristics of participants stratified by TB infection status**")
```


```{r, echo=FALSE}
knitr::kable(table2, row.names = FALSE, caption = "**Univariate and multivariate analysis**")
```


```{r Figure 1, echo=FALSE}

```

